---
title: "QMM Assignment 2"
author: "Akash Nair"
date: "2024-09-22"
output: html_document
---
The Weigelt Corporation has three branch plants with excess production capacity. Fortunately,
the corporation has a new product ready to begin production, and all three plants have this
capability, so some of the excess capacity can be used in this way. This product can be made in
three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300,
respectively. Plants 1, 2, and 3 have the excess capacity to produce 750, 900, and 450 units per
day of this product, respectively, regardless of the size or combination of sizes involved.
The amount of available in-process storage space also imposes a limitation on the production
rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet,
respectively, of in-process storage space available for a day's production of this product. Each
unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet,
respectively.
Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and
small sizes, respectively, would be sold per day.
At each plant, some employees will need to be laid off unless most of the plant’s excess
production capacity can be used to produce the new product. To avoid layoffs if possible,
management has decided that the plants should use the same percentage of their excess capacity
to produce the new product.
Management wishes to know how much of each of the sizes should be produced by each of the
plants to maximize profit.

Solution:

```{r}
data = matrix(c("a_l", "a_m", "a_s", "b_l", "b_m", "b_s", "c_l", "c_m", "c_s"), ncol=3, byrow=TRUE)

colnames(data) = c('Large', 'Medium', 'Small')
rownames(data) = c("Plant 1", "Plant 2", "Plant 3")

final=as.table(data)

final
```
Lets assume that

Products produced by Plant 1 are represented by: 
$$al, am, as$$

Products produced by Plant 2 are represented by:
$$bl, bm, bs$$


Products produced by Plant 3 are represented by:
$$cl, cm, cs$$

Let us consider “p” as production capacity percentage.

The above products are in the order of Large, Medium and Small Respectively
(a)The decision variables are 
$$al, am, as, bl, bm, bs, cl, cm, cs$$ 
(b) Formulating a linear programming model

$$Maximize Z = (420al+ 360am+ 300as) + (420bl+ 360bm+ 300bs) + (420cl+ 360cm+ 300cs)$$
#Constraints

1) Storage Space Capacity constraints:

Plant 1: $$20al + 15am + 12as \le 13000$$

Plant 2: $$20bl + 15bm + 12bs \le 12000$$

Plant 3: $$20cl + 15cm + 12cs \le 5000$$

2) Production capacity constraints:

Given Plant 1, Plant 2 and Plant 3 have the capacity to produce 750, 900 and 450 units per day of this product, respectively, regardless of size or combinations of sizes

Plant 1: $$al + am + as \le 750$$
Plant 2: $$bl + bm + bs \le 900$$
Plant 3: $$cl + cm + cs \le 450$$

3) Sales Constraints:

The total production of each size (Large, Medium, small) should not exceed the given limit across all plants

Large:  $$al + bl + cl \le 900$$
Medium: $$am + bm + cm \le 1200$$
Small:  $$as + bs + cs \le 750$$

4) Same Percentage capacity constraint:

Since the management has decided to use the same percentage of excess capacity of all plants to avoid playoffs, it can be expressed as follows-

$$P = \frac{al + am + as}{750} = \frac{bl + bm + bs}{900} = \frac{cl + cm + cs}{450}$$
5) Non- Negativity Constraints : 
$$al \geq 0, \; am \geq 0, \; as \geq 0, \; bl \geq 0, \; bm \geq 0, \; bs \geq 0, \; cl \geq 0, \; cm \geq 0, \; cs \geq 0$$
Solving the LP Problem using ‘lpSolve’ function 


Step 1: Install and load the ‘lpSolve’ Package.


```{r}
#install.packages("lpSolve")
```
```{r}
library(lpSolve)
```

Step 2: Prepare a vector of the objective function

#Objective function vector variables
```{r}
f.obj = c(420, 360, 300, 420, 360, 300, 420, 360, 300)
```

step 3:

#Constraint matrix with variables
```{r}
f.con = matrix(c(20, 15, 12, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 20, 15, 12, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 20, 15, 12,
                 1, 1, 1, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 1, 1, 1, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 1, 1, 1,
                 1, 0, 0, 1, 0, 0, 1, 0, 0,
                 0, 1, 0, 0, 1, 0, 0, 1, 0,
                 0, 0, 1, 0, 0, 1, 0, 0, 1,
                 900, 900, 900, -750, -750, -750, 0, 0, 0,
                 0, 0, 0, 450, 450, 450, -900, -900, -900,
                 450, 450, 450, 0, 0, 0, -750, -750, -750
                 ), nrow = 12, ncol = 9, byrow = TRUE)


f.con

```

Step 4:

#Creating Constraint directions vector
```{r}
f.dir <- c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=")
```

Step 5: 

#vector for constraints
```{r}
f.rhs <- c(13000, 12000, 5000, 750, 900, 450, 900, 1200, 750, 0, 0, 0)
```

```{r}
lp("max", f.obj, f.con, f.dir, f.rhs)

```

#Displaying the solution
```{r}
lp('max', f.obj, f.con, f.dir, f.rhs)$solution
```
This command retrieves the optimal values for the decision variables (quantities of products to produce at each plant).

# Interpretation of the solution

Plant 1:
Large products: 516.67 units
Medium products: 177.78 units
Small products: 0 units

Plant 2:
Large products: 0 units
Medium products: 666.67 units
Small products: 166.67 units

Plant 3:
Large products: 0 units
Medium products: 0 units
Small products: 416.67 units

# Conclusion
This solution shows how many of each product should be made at each plant to get the most profit while following the set rules. With this production plan, the highest profit possible is 696,000. This method helps use resources effectively in production.


